Section Three

### 2020-10-04 new2020/sections/sec03.Rmd
texTalpha pkg
Try find the bug

---
title: Section Three
header-includes:
   - \usepackage{cleveref}
   - \input{../symbols.tex}
---

@@

The overall values are shown in @cref:tab:results1;. The particle filter predictions are slightly more accurate than GTFS on average, and have about 20% error versus GTFS's 27%. However, only 77% of prediction intervals contain the actual arrival time, indicating that uncertainties are being under-predicted (the nominal coverage is 85%).

Since traffic and bus numbers change over the course of a day, we also computed the assessment statistics over 15\ minute intervals. These results are shown in @cref:fig:results2;, and immediately we can see the sizeable affect of peak time of predictive accuracy. This suggests that, during off-peak time when traffic conditions are stable, the real-time predictions are a vast improvment on the status quo. Since the underlying network model is not yet enabled to handle forecasts, its not surprising that accuracy and uncertainty coverage is lacking.

```{r day_results,echo=FALSE,fig.width=8,fig.height=4,message=FALSE,warning=FALSE,cache=TRUE,fig.cap="Prediction accuracy results for the particle filter and GTFS methods."}
# library(RSQLite)
library(tidyverse)
# library(dbplyr)
library(patchwork)

load("../data/etadata.rda")
res_timeofday <- eta_talk %>%
    mutate(
        hour = timestamp %>% format("%H") %>% as.integer,
        minute = timestamp %>% format("%M") %>% as.integer,
        hour = hour + (floor(minute / 15) * 15 / 60)
    ) %>%
    filter(time_until_arrival > 0) %>%
    group_by(hour) %>%
    summarize(
        pf_rmse = sqrt(mean(pf_error^2)),
        pf_mae = mean(abs(pf_error)),
        pf_mape = mean(abs(pf_error / time_until_arrival)),
        pf_ci_cov = mean(pf_ci),
        pf_ci_width = mean(pf_upper - pf_lower),
        pf_ci_lower = quantile(pf_upper - pf_lower, 0.025),
        pf_ci_upper = quantile(pf_upper - pf_lower, 0.975),
        gtfs_rmse = sqrt(mean(gtfs_error^2)),
        gtfs_mae = mean(abs(gtfs_error)),
        gtfs_mape = mean(abs(gtfs_error / time_until_arrival))
    ) %>% arrange(hour)

pal <- viridis::magma(3)[2:1] #RColorBrewer::brewer.pal(2, "Set1")
names(pal) <- c("Particle filter", "GTFS")

plot_base <- ggplot(res_timeofday, aes(hour)) +
    theme_classic() +
    theme(legend.position = "none") +
    scale_colour_manual(values = pal, name = "") +
    xlab("Time of day (hour)")

plot_rmse <- plot_base +
    geom_path(aes(y = pf_rmse, colour = "Particle filter")) +
    geom_path(aes(y = gtfs_rmse, colour = "GTFS"), alpha = 0.5) +
    scale_y_continuous(
        "RMSE (s)"
        # trans = "log10"
    ) +
    theme(legend.position = "bottom")

plot_mae <- plot_base +
    geom_path(aes(y = pf_mae, colour = "Particle filter")) +
    geom_path(aes(y = gtfs_mae, colour = "GTFS"), alpha = 0.5) +
    scale_y_continuous(
        "MAE (s)"
        # trans = "log10"
    )

plot_mape <- plot_base +
    geom_path(aes(y = 100 * pf_mape, colour = "Particle filter")) +
    geom_path(aes(y = 100 * gtfs_mape, colour = "GTFS"), alpha = 0.5) +
    scale_y_continuous(
        "MAPE (%)"
        # trans = "log10"
    )

plot_ci <- plot_base +
    geom_hline(aes(yintercept = 85), lty = 2) +
    geom_path(aes(y = 100 * pf_ci_cov, colour = "Particle filter")) +
    scale_y_continuous(
        name = "PICP (%)"
    )

plot_rmse + plot_mae +
    plot_mape + plot_ci +
    guide_area() +
    plot_layout(
        guides = "collect",
        design = "
            13
            24
            55
        ",
        heights = c(1, 1, 0.1)
    )
```

>>>>

---
title: Section Three

header-includes:
   - \usepackage{cleveref}
   - \usepackage{booktabs}
---

@@

The overall values are shown in @cref:tab:overall_results;. The particle filter predictions are slightly more accurate than GTFS on average, and have about 20% error versus GTFS's 27%. However, only 77% of prediction intervals contain the actual arrival time, indicating that uncertainties are being under-predicted (the nominal coverage is 85%).

```{r overall_results,echo=FALSE,cache=TRUE,results="asis",dependson="load_results"}
# res_overall <- eta_talk %>%
#    filter(time_until_arrival > 0) %>%
#    summarize(
#       pf_rmse = sqrt(mean(pf_error^2)),
#       pf_mae = mean(abs(pf_error)),
#       pf_mape = mean(abs(pf_error / time_until_arrival)) * 100,
#       pf_cicov = mean(pf_ci) * 100,
#    #   pf_ci_width = mean(pf_upper - pf_lower),
#    #   pf_ci_lower = quantile(pf_upper - pf_lower, 0.025),
#    #   pf_ci_upper = quantile(pf_upper - pf_lower, 0.975),
#       gtfs_rmse = sqrt(mean(gtfs_error^2)),
#       gtfs_mae = mean(abs(gtfs_error)),
#       gtfs_mape = mean(abs(gtfs_error / time_until_arrival)) * 100
#    ) %>%
#    pivot_longer(everything(),
#       names_to = c("model", ".value"),
#       names_pattern = "([a-z]*)_(.*)"
#    ) %>%
#    mutate(
#       model = c("Particle filter", "Schedule-delay")
#    ) %>%
#    rename(
#       " " = "model",
#       "RMSE (s)" = "rmse",
#       "MAE (s)" = "mae",
#       "MAPE (%)" = "mape",
#       "PICP (%)" = "cicov"
#    )

# opts <- options(knitr.kable.NA = "")
# knitr::kable(res_overall, #"latex",
#    booktabs = TRUE,
#    # table.envir = "table",
#    digits = 0,
#    caption = "Results for particle filter and other."
# )
# options(opts)
```

Since traffic and bus numbers change over the course of a day, we also computed the assessment statistics over 15\ minute intervals. These results are shown in @cref:fig:day_results;, and immediately we can see the sizeable affect of peak time of predictive accuracy. This suggests that, during off-peak time when traffic conditions are stable, the real-time predictions are a vast improvement on the status quo. Since the underlying network model is not yet enabled to handle forecasts, its not surprising that accuracy and uncertainty coverage is lacking.

```{r day_results,echo=FALSE,fig.width=8,fig.height=4,message=FALSE,warning=FALSE,cache=TRUE,fig.cap="\\label{fig:day_results}Prediction accuracy results for the particle filter and GTFS methods.",dependson="load_results"}
# library(patchwork)

# res_timeofday <- eta_talk %>%
#     mutate(
#         hour = timestamp %>% format("%H") %>% as.integer,
#         minute = timestamp %>% format("%M") %>% as.integer,
#         hour = hour + (floor(minute / 15) * 15 / 60)
#     ) %>%
#     filter(time_until_arrival > 0) %>%
#     group_by(hour) %>%
#     summarize(
#         pf_rmse = sqrt(mean(pf_error^2)),
#         pf_mae = mean(abs(pf_error)),
#         pf_mape = mean(abs(pf_error / time_until_arrival)),
#         pf_ci_cov = mean(pf_ci),
#         pf_ci_width = mean(pf_upper - pf_lower),
#         pf_ci_lower = quantile(pf_upper - pf_lower, 0.025),
#         pf_ci_upper = quantile(pf_upper - pf_lower, 0.975),
#         gtfs_rmse = sqrt(mean(gtfs_error^2)),
#         gtfs_mae = mean(abs(gtfs_error)),
#         gtfs_mape = mean(abs(gtfs_error / time_until_arrival))
#     ) %>% arrange(hour)

# pal <- viridis::magma(3)[2:1] #RColorBrewer::brewer.pal(2, "Set1")
# names(pal) <- c("Particle filter", "GTFS")

# plot_base <- ggplot(res_timeofday, aes(hour)) +
#     theme_classic() +
#     theme(legend.position = "none") +
#     scale_colour_manual(values = pal, name = "") +
#     xlab("Time of day (hour)")

# plot_rmse <- plot_base +
#     geom_path(aes(y = pf_rmse, colour = "Particle filter")) +
#     geom_path(aes(y = gtfs_rmse, colour = "GTFS"), alpha = 0.5) +
#     scale_y_continuous(
#         "RMSE (s)"
#         # trans = "log10"
#     ) +
#     theme(legend.position = "bottom")

# plot_mae <- plot_base +
#     geom_path(aes(y = pf_mae, colour = "Particle filter")) +
#     geom_path(aes(y = gtfs_mae, colour = "GTFS"), alpha = 0.5) +
#     scale_y_continuous(
#         "MAE (s)"
#         # trans = "log10"
#     )

# plot_mape <- plot_base +
#     geom_path(aes(y = 100 * pf_mape, colour = "Particle filter")) +
#     geom_path(aes(y = 100 * gtfs_mape, colour = "GTFS"), alpha = 0.5) +
#     scale_y_continuous(
#         "MAPE (%)"
#         # trans = "log10"
#     )

# plot_ci <- plot_base +
#     geom_hline(aes(yintercept = 85), lty = 2) +
#     geom_path(aes(y = 100 * pf_ci_cov, colour = "Particle filter")) +
#     scale_y_continuous(
#         name = "PICP (%)"
#     )

# plot_rmse + plot_mae +
#     plot_mape + plot_ci +
#     guide_area() +
#     plot_layout(
#         guides = "collect",
#         design = "
#             13
#             24
#             55
#         ",
#         heights = c(1, 1, 0.1)
#     )
```
